---
<<<<<<< HEAD
title: "DATA VISUALIZATION AND REGRESSION"
output: html_document
---

```{r}
install.packages("mnormt")
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Bivariate OLS regression
```{r}
f1 <- lm(attend_school ~ banks, data = master_2002_banks)
summary(f1) 
=======
title: "DATA VISUALIZATION"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
View(master_2002_banks)
```

```{r}
my_plot <- ggplot(master_2002,aes(speed,dist)) + geom_point()
my_plot
```





## Bivariate OLS regression
```{r}
f1 <- lm(attend_school ~ banks, data = master_2002_banks)
summary(f1)
>>>>>>> d35fc7ea782e29e366c8104db2a06dd70722b5a3
```

```{r}
library(ggplot2)
```

```{r}
<<<<<<< HEAD
ggplot(master_2002_banks, aes(x = banks, y = attend_school)) +  
  geom_point() + 
  stat_smooth(method = "lm", col = "red")
```


```{r}
age <- as.numeric(master_2002_banks$age)
```

## OLS regression with control variables 
```{r}
f2 <- lm(attend_school ~ banks + age + hours_sleep + hh_avg_ed + hh_income + cult_act + domestic_hwork + factor(municipality), data = master_2002_banks) 
summary(f2) 
# should i add  + factor(municipality) here?
```

## Linear regression allowing for the effect of banks to be different by age (convert age to dummies and create the proper interactions)
=======
ggplot(master_2002_banks, aes(x = banks, y = attend_school)) + 
  geom_point() +
  stat_smooth(method = "lm", col = "red")
```

## OLS regression with control variables 
```{r}
f2 <- lm(attend_school ~ banks + age + hh_avg_ed + hh_income + cult_act + domestic_hwork, data = master_2002_banks)
summary(f2)
```

## Linear regression allowing for the effect of banks to be different by age (convert age to dummies and create the proper interactions)


>>>>>>> d35fc7ea782e29e366c8104db2a06dd70722b5a3
```{r}
install.packages("sjPlot")
install.packages("sjmisc")
```

<<<<<<< HEAD
```{r}
library(sjPlot)
library(sjmisc)

# fit model with interaction
f3 <- lm(attend_school ~ banks + age + hours_sleep + hh_avg_ed + hh_income + age + cult_act + domestic_hwork + age*hh_income, data = master_2002_banks)
summary(f3)
plot_model(f3, type = "pred", terms = c("hh_income", "age"))
```

=======

```{r}
library(sjPlot)
library(sjmisc)
data(master_2002_banks)

# fit model with interaction
f3 <- lm(attend_school ~ banks + hh_avg_ed + hh_income + cult_act + domestic_hwork + age*hh_income, data = master_2002_banks)

summary(f3)

plot_model(f3, type = "pred", terms = c("hh_income", "age"))
```


```{r}
f4 <- lm(attend_school ~ banks*age*hh_income + hh_avg_ed + cult_act + domestic_hwork, data = master_2002_banks)

summary(f4)

plot_model(f4, type = "pred", terms = c("hh_income", "age", "banks"))
```


>>>>>>> d35fc7ea782e29e366c8104db2a06dd70722b5a3
## logit regression
```{r}
install.packages("margins")
install.packages("ggeffects")
```

```{r}
library(ggeffects)
library("lme4")
```

<<<<<<< HEAD
=======

>>>>>>> d35fc7ea782e29e366c8104db2a06dd70722b5a3
```{r}
f5 <- glmer(attend_school ~ banks + (1|municipality), data = master_2002_banks, family = 'binomial')
print(f5, corr = FALSE)
summary(f5)
```
<<<<<<< HEAD

=======
>>>>>>> d35fc7ea782e29e366c8104db2a06dd70722b5a3
```{r}
plot(fitted(f5), residuals(f5), xlab = "Fitted Values", ylab = "Residuals")
abline(h = 0, lty = 2)
```


```{r}
<<<<<<< HEAD
f6 <- glmer(attend_school ~ banks + age + hours_sleep + hh_avg_ed + hh_income + cult_act + domestic_hwork + (1|municipality), data = master_2002_banks, family = 'binomial')
print(f6, corr = FALSE)
```

=======
f6 <- glmer(attend_school ~ banks + age + hh_avg_ed + hh_income + cult_act + domestic_hwork + (1|municipality), data = master_2002_banks, family = 'binomial')
print(f5, corr = FALSE)
```

## turn age into numeric

>>>>>>> d35fc7ea782e29e366c8104db2a06dd70722b5a3
```{r}
plot(fitted(f6), residuals(f6), xlab = "Fitted Values", ylab = "Residuals")
abline(h = 0, lty = 2)
```

```{r}
<<<<<<< HEAD
margin_eff1 <- ggpredict(f6, term = "banks")
plot(margin_eff1)
```

```{r}
margin_eff2 <- ggpredict(f6, term = "hh_income")
plot(margin_eff2)
```

```{r}
margin_eff3 <- ggpredict(f6, term = "age")
plot(margin_eff3)
```

```{r}
margin_eff4 <- ggpredict(f6, term = "hh_avg_ed")
plot(margin_eff4)
```

```{r}
margin_eff5 <- ggpredict(f6, term = "hours_sleep")
plot(margin_eff5)
```
=======
library(MuMIn)
```

```{r}
r.squaredGLMM(f5)
```

```{r}
margin_eff <- ggpredict(f5, term = "banks")
margin_eff
```

```{r}
plot(margin_eff)
```


>>>>>>> d35fc7ea782e29e366c8104db2a06dd70722b5a3
